---
layout: post
title: "Relationship Analysis"
author: "Paulin"
date: "2015-12-25"
output: md_document
categories: R
---
```{r, echo = FALSE, eval = FALSE}
# TODO: Turn this into a function that will handle the world!!!!!
#' @description a function that will handle factors
#' @param x a vector of factors
#' @param oldStatusLevels the levels to be replaced # TODO: change this into a named list
#' @param keepStatusLevels a vector of the factors to keep the same
#' @param removeStatusLevels an optional vector of the factors to remove
tidyStatus <- function(x){
  oldStatusLevels <- c("Partner deceased", "Still together") # the levels to be replaced
  keepStatusLevels <- "Broke up"
  newStatusLevel <- "Broke up or partner deceased" # must be single character
  removeStatusLevels <- NULL
  levels(x) <- c(sort(c(oldStatusLevels, keepStatusLevels)), newStatusLevel)
  res <- ifelse(x %in% oldStatusLevels, newStatusLevel, x)
  levels(res) <- c(sort(keepStatusLevels), newStatusLevel)
  return(res)
}


```
```{r, echo = FALSE, eval = TRUE}
## General prep
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, error = FALSE)

# set default chunk options


# source: http://data.stanford.edu/hcmst
# translate .Rmd files in the _sources dir to .md files in _posts
# if jekyll is installed then it also compiles the site in _site
# use serve = FALSE if you do not have jekyll installed/ don't want to preview the site
# setwd("/home/paulin/Documents/paulinshek.github.io")
# servr::jekyll() 

# ggplot2 theme
library(ggplot2)
# TODO
```

```{r}
## Data prep

# Navigate to where the data lives. Do not put it on github
dataPath <- "/home/paulin/Documents/SomeRLovin/Relationship Prediction/data/HCMST"

# convert the data from sav files (SAS) to RData objects
library(foreign)

# How Couples Stay Together data
hcmst <- as.data.frame(read.spss(file.path(dataPath, "HCMST.sav"))) # waves 1,2,3
hcmst4 <- as.data.frame(read.spss(file.path(dataPath, "HCMSTwave4.sav")))
hcmst5 <- as.data.frame(read.spss(file.path(dataPath, "HCMSTwave5.sav")))

```

Joe and I are both data fanatics, and we have been thinking of working on a project together for a long time. When we first started out, I used to always joke that we should do some analysis on other peoples relationships and use it to try and predict how long we'd last. Now we are over two years into our relationship, and we've finally deemed it stable enough to withhold the conclusions that we might find from the data! So here we go...

THE DATA:
We used the HCMST dataset, which is a comprehensive dataset of survey data, featuring over 4000 people. It features 5 "waves" of surveys, spread over 5 years. For each person in the survey, their partner may not necessarily be also included in the survey, but there are ... many couples in the survey. We deemed this a sufficiently high number to only look at complete couples interviewed in the survey. 



```{r, fig.cap="Grey denotes the repondents who declined to answer a follow-up survey, or where not in a relationship in wave 1, in the first place", warning=FALSE}
# TODO: make a dataset which joins the waves together, keeping the columns related to break up at each wave
library(dplyr)
relationshipStatusOverWave <- data.frame(`Original relationship status` = hcmst$qflag, 
                           `Status at wave 2` = hcmst$w2_broke_up, 
                           `Status at wave 3` = hcmst$w3_broke_up, 
                           `Status at wave 4` = hcmst4$w4_broke_up, 
                           `Status at wave 5` = hcmst5$w5_broke_up,
                           check.names = FALSE)

relationshipStatusOverWave <- relationshipStatusOverWave %>% 
  filter_(~`Original relationship status`=="partnered") %>% 
  select(starts_with("Status"))


tempCombine <- function(x){
  x <- as.character(x)
  res <- ifelse(x %in% c("broke up", "partner deceased", "partner passed away"), "Broke up or partner deceased", x)
  res <- ifelse(res=="still together", "Still together", res)
  levels(res) <- c("Still together", "Broke up or partner deceased")
  return(res)
}
relationshipStatusOverWave <- lapply(relationshipStatusOverWave, tempCombine)

N <- length(relationshipStatusOverWave[[1]])
relationshipStatusOverWave <- lapply(relationshipStatusOverWave, as.character)
status <- do.call(c, relationshipStatusOverWave)
names(status) <- NULL
status <- factor(status, levels = c("Still together", "Broke up or partner deceased"))
wave <- rep(2:5, each = N)
longData <- data.frame(Wave = wave, Status = status)


library(ggplot2)
ggplot(data = longData, aes(x = Wave, fill = Status)) + geom_bar() + 
  ggtitle("Barchart to show the relationship status \nof each survey respondent at each wave \nof the survey") + ylab("Count")

```